---
title: "Chapters 8 and 9 Homework"
author: "Reid Ginoza"
date: "11/5/2019"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(pander)
```

```{r Support Functions}
### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits = 4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}
```

# Chapters 8 and 9 Homework

Data is from Problem 8.29.
Coloration (on a scale of 1-10) of strawberries grouped by 4 different preservatives.

```{r Import Strawberry Discoloration Data}
discolor_data <- as_tibble(read_csv("ex8-29.csv", skip = 1, col_names = c("Group I", "Group II", "Group III", "Group IV"), col_types = "dddd")) %>% gather(key = "Preservative", value = "Discoloration")

```

## 1.  One-Way ANOVA
### (a)  Find the means and standard deviations of the groups.

```{r Mean and SD}
discolor_summary_1 <- discolor_data %>% group_by(Preservative) %>% summarise(Mean = mean(Discoloration), SD = round(sd(Discoloration), digits = 3))
discolor_summary_1 %>% select(Preservative, Mean, SD) %>% kable(align ='lrr',format.args = list(trim = TRUE)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                font_size = 14,
                position = "left")
```

### (b)  Construct the corresponding one-way ANOVA table.
```{r one way ANOVA table}
(aov(Discoloration ~ Preservative, data=discolor_data) %>% summary() -> discoloration_one.way.table) %>% pander(style='rmarkdown')
```

### (c)  Formally test to determine if there is a difference between the groups at the $\alpha = 0.05$ level.
**Hypotheses**<br>
$H_0$: $\mu_{I} = \mu_{II} = \mu_{III} = \mu_{IV}$<br>
$H_1$: at least one is different

**Test Statistic**<br>
$F_0 = `r discoloration_one.way.table[[1]][["F value"]][1]`$

***p*-value**<br>
$`r p.value.string(discoloration_one.way.table[[1]][["Pr(>F)"]][1])`$.

**Rejection Region**<br>
Reject $H_0$ if $p < \alpha$, where $\alpha=0.05$.

**Conclusion**<br>
Reject $H_0$. There is sufficient evidence to suggest that at least one preservative is different from the others.
