---
title: "Chapters 8 and 9 Homework"
author: "Reid Ginoza"
date: "11/5/2019"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(pander)
```

```{r Support Functions}
### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits = 4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}
```

# Chapters 8 and 9 Homework

Data is from Problem 8.29.
Coloration (on a scale of 1-10) of strawberries grouped by 4 different preservatives.

```{r Import Strawberry Discoloration Data}
discolor_data <- as_tibble(read_csv("ex8-29.csv", skip = 1, col_names = c("Group I", "Group II", "Group III", "Group IV"), col_types = "dddd")) %>% gather(key = "Preservative", value = "Discoloration")

```


## 2.  Kruskal-Wallis
### (a)  Find the median and IQR of each group.
```{r Median and IQR}
discolor_summary_2 <- discolor_data %>% group_by(Preservative) %>% summarise(Median = median(Discoloration), IQR = round(IQR(Discoloration), digits = 3))
discolor_summary_2 %>% select(Preservative, Median, IQR) %>% kable(align ='lrr',format.args = list(trim = TRUE)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                font_size = 14,
                position = "left")
```

### (b)  Rank the data as needed for the Kruskal-Wallis test.  Display your ranks using a table.
I'm displaying the table with rankings in the original sorting on the left and sorted by rank on the right.
```{r Kruskal-Wallis Ranking}
# Want to list whether a rank is in a tie.
# Use with lapply
is.tied = function(x){
  if (x==TRUE) {
    return("__tied")
  } else {
    return("")
  }
}

discolor_data$Ranking <- rank(discolor_data$Discoloration, na.last = "na", ties.method = "average")
discolor_data$Tied <- lapply(Reduce("|",tibble(duplicated(discolor_data$Ranking),duplicated(discolor_data$Ranking,fromLast = TRUE))),is.tied)

discolor_data %>% select(Preservative, Discoloration, Ranking, Tied) %>% 
  arrange(Ranking, Preservative) -> discolor_rank.sorted
discolor_data %>% select(Preservative, Discoloration, Ranking, Tied) -> discolor_original.sorted
kable(list(discolor_original.sorted, discolor_rank.sorted)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                font_size = 14,
                position = "left")
```

### (c)  Formally test for a difference between the groups using the Kruskal-Wallis test at the $\alpha = 0.05$ level.

```{r Kruskal-Wallis Hypothesis Test}
discolor_kw_h.test <- kruskal.test(Discoloration ~ Preservative, data = discolor_data)
```

**Hypotheses**<br>
$H_0: \ \textsf{M}_{I} = \textsf{M}_{II} = \textsf{M}_{III} = \textsf{M}_{IV}$ <br> 
$H_1:$ at least one is different


**Test Statistic**<br>
$H = `r round(discolor_kw_h.test$statistic,digits=2)`$.


***p*-value**<br>
$`r p.value.string(discolor_kw_h.test$p.value)`$.

**Conclusion**<br>
Reject $H_0$. There is sufficient evidence to suggest that at least one of the preservatives is different from the others.
