---
title: "Chpt 6 Part 3"
author: "Reid Ginoza"
date: "9/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stats)
library(infer)
library(groupedstats)
library(knitr)
library(kableExtra)
```

```{r Support Functions}
### Support Functions ###

# Makes the ranking look nice. Returns integers unless there's a decimal.
# Will be applied element-wise using lapply(). (Not a vector function ie. with apply())
rank_round = function(x){
  if (x %% 1 == 0) {
    return(round(x, digits = 0))
  } else {
    return(round(x, digits = 1))
  }
}

### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits=4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}
```

## 3. Data from problem 6.57 (page 359)
Soil pH levels at 15 locations before and after mining.
```{r Problem 3 load data}
soil_ph_data <- as_tibble(read_csv("6_57.csv", col_names = TRUE, col_types = 'idd'))
```

### (a) Is this independent or paired data? Why?
This data is paired since the same location is measured twice, once in the "Before" group and once in the "After" group.

### (b) Which $t$-test would be appropriate to compare the two groups?
Since the data is paired, we will use the paired $t$-test.

### (c) Find the mean and standard deviation of the data, either split by before and after mining or as the difference between before and after mining.
In the paired $t$-test, we use the difference between the before and after measurements for each location.

```{r Soil Difference between Before and After}
soil_ph_data$Difference <- soil_ph_data$Before - soil_ph_data$After
soil_diff.mean <- round(mean(soil_ph_data$Difference), digits = 2)
soil_diff.sd <- round(sd(soil_ph_data$Difference), digits = 2)
```

The mean (sd) of the difference between before and after mining pH level is `r soil_diff.mean` (`r soil_diff.sd`).

### (d) Formally test to determine if the mean pH level is different after strip-mining. Use the appropriate version of the t-test at the α = 0.05 level.

```{r Soil t-test}
soil_t.test <- t.test(soil_ph_data$Before, soil_ph_data$After, paired = TRUE, alternative = "two.sided", conf.level = 0.95)
# soil_ph_data %>% pull(Difference) %>% t.test(mu=0, alternative="two.sided") -> soil_t.test2 # using pipe operator

soil_t.test$p.value %>% p.value.string -> soil_p.value
```

**Hypotheses**  
$H_0: \mu_a = \mu_b$
$H_1: \mu_a \ne \mu_b$

**Test Statistic**  
$t_0 = `r soil_t.test$statistic`$  
$`r soil_p.value`$

**Rejection Region**  
Reject $H_0$ if $p<\alpha$, where $\alpha=0.05$.

**Conclusion**  
Reject $H_0$. There is sufficient evidence to suggest that the strip-mining for coal changed the pH level of the soil.

### (e) Construct the 95% confidence interval for µB − µA.
The 95% confidence interval for the difference of mean pH level is (`r soil_t.test$conf.int`).

### (f) Is the t-test appropriate here? Justify your answer.
A paired $t$-test requires the two samples to be approximately normal. Here is the qq-plot for the difference between the two groups:

```{r Soil Before and After qqplot}
soil.before_qq.plot <- soil_ph_data %>% ggplot(aes(sample = Before)) +
        stat_qq() + 
        stat_qq_line() +
        theme_bw() +
        ggtitle('Quantile-Quantile Plot of pH Level Before')
soil.before_qq.plot

soil.after_qq.plot <- soil_ph_data %>% ggplot(aes(sample = After)) +
        stat_qq() + 
        stat_qq_line() +
        theme_bw() +
        ggtitle('Quantile-Quantile Plot of pH Level After')
soil.after_qq.plot
```

For both samples, the data points fall near the line, except at the tails, and the samples are approximately normal. This then implies that the difference between the two samples are approximately normal, which we can also visually inspect with a qq plot of the difference. The difference is also approximately normal, so the normality assumptions are satisfied for the paired $t$-test.

```{r Soil Difference qqplot}
soil.difference_qq.plot <- soil_ph_data %>% ggplot(aes(sample = Difference)) +
        stat_qq() + 
        stat_qq_line() +
        theme_bw() +
        ggtitle('Quantile-Quantile Plot of pH Level Difference')
soil.difference_qq.plot
```

The samples from the different locations are also independent. So the $t$-test is an appropriate test. Also, the sample size is small, $n<30$, so a $z$-test on the difference would not make sense.
