---
title: "Stats Modeling Project"
author: 'Group 5: Xin Jin, Reid Ginoza, Heidi Lovejoy'
date: "11/18/2019"
output:
  html_document:
    df_print: paged
nocite: |
  @rlang
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra) # for tables
library(magrittr)  # provides more pipe oeprators
library(haven)
options(knitr.table.format = "latex")
```

```{r import data}
ncaa.div.i_data <- read_spss('InstLevel.sav') %>%
  select(  # TODO will change later when we decide what variables we want
  -unitid, -addr1_txt, -addr2_txt, -city_txt, -zip_text, -sector_cd,
  -ClassificationCode, -ClassificationOther
  ) %>%
  filter(grepl('4-year', sector_name))  %>% # only 4-year Schools
  filter(is.na(IL_PARTIC_COED_MEN) | IL_PARTIC_COED_MEN == 0 ) %>% # only schools with no male participants in coed
  filter(is.na(IL_PARTIC_COED_WOMEN) | IL_PARTIC_COED_WOMEN == 0 ) %>% # only schools with no female participants in coed
  select(-contains("COED")) %>% # ignore variables with the word "coed"
  filter(grepl('NCAA Division I-', classification_name)) # only NCAA division 1
```

```{r Notes to self, include=FALSE}
##### All 4-year Institutions
# institution_data <- read_spss('InstLevel.sav') %>%
#   select(  # TODO will change later when we decide what variables we want
#   -unitid, -addr1_txt, -addr2_txt, -city_txt, -zip_text, -sector_cd,
#   -ClassificationCode, -ClassificationOther
#   ) %>%
#   filter(grepl('4-year', sector_name))  %>% # only 4-year Schools
#   filter(is.na(IL_PARTIC_COED_MEN) | IL_PARTIC_COED_MEN == 0 ) %>% # only schools with no male participants in coed
#   filter(is.na(IL_PARTIC_COED_WOMEN) | IL_PARTIC_COED_WOMEN == 0 ) %>% # only schools with no female participants in coed
#   select(-contains("COED")) # ignore variables with the word "coed"
##### To view the school data
# school_data <- read_spss('Schools.sav') %>% select(
#   -unitid, -addr1_txt, -addr2_txt, -zip_text, -ClassificationCode,-ClassificationOther,
#   -sector_cd
#   )
#####
# To view variables included as a list (although you could probably just view the data):
# institution_columns <- institution_data %>% colnames()
#####
# To view number of NAs in a column:
# count.of.na <- institution_data %>% summarise_all(~sum(is.na(.)))
```

# Introduction
<!-- %TODO Briefly give meaning to your project – what is the background and why is this project important?  (20 points) -->

This project examines potential(?)salary differences between head coaches from public and private NCAA Division I institutions. We used the data from the Equity in Athletics Survey, Year 2017-2018, from the U.S. Department of Education Office of Postsecondary Education [-@EADA2018]. 

# Hypotheses

## Head Coaches of Men's Teams and School Type
<!-- %TODO Clearly state the hypotheses being tested. (20 points) -->
<!-- Note: we need the two spaces at the end of the first line of the hypothesis test-->

## Hypothesis 1) Formally test to determine if there is a difference in the variances of head coach salaries in public institutions vs. those in private institutions. 

## Method 1) To compare the population variances of two independent, randomly selected samples of size n_1 and n_2 (NOTE: fill these in with actual numbers), we use the F-distribution method.

<!-- Note, this requires normality, but the textbook example specifically states that it ignores a slight skew, so this is something to keep in mind-->

<!-- Note:  Just typing out the skeleton code for variance testing. It didn't look like Reid already named things, so I put in the placeholders to start with below-->

```{r}
#salary_var <- var.test(public, private, alternative = "two.sided", conf.level = 0.95, ratio = 1)

#salary_p <- p.value.string(salary_var$p.value)
```

**Hypotheses**

$H_0: \ \sigma^2_{public} = \sigma^2_{private}$ <br> 
$H_1: \ \sigma^2_{public} \ne \sigma^2_{private}$

**Test Statistic**

$F_0 = `r round(salary_var$statistic, digits=2)`$

***p*-value**

$`r salary_p`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Reject/FTR $H_0$. There is/is not sufficient evidence to suggest that there is a difference in the variances of head coach salaries in public institutions vs. those in private institutions.  

## Hypothesis 2) Formally test to determine if there is a difference in the means of head coach salaries in public institutions vs. those in private institutions. 

## Method 2) To compare the population means of the two populations, we use...

<!--TODO: This test will, of course, depend on A) the results from variance test, and B) a test for normality (Q-Q is best)-->

<!--EQUAL VARIANCE/NOT NORMAL:  Wilcoxon Rank Sum-->

<!--UNEQUAL VARIANCES/NORMAL: t-Test/Satterthwaites-->

<!--UNEQUAL VARIANCE/NOT NORMAL: I'm a bit confused about this part. Ch. 5 discusses this, read: large-sample approximation w/unequal variances. We didn't go over this in class, though, so I feel like I may be missing something or just not thinking clearly. In any case, we'll figure this out when we get there-->

## Hypothesis 3)  

## Method 3) 


## Hypothesis 4) 

## Method 4) 


## Hypothesis 5)  

## Method 5) 

# Results
<!-- Describe the data (1st paragraph) as well as analysis results (order of results should correspond to hypotheses listed)-->

<!--Describe the data:-->

The data analyzed comes from the Equity in Athletics Survey, Year 2017-2018, from the U.S. Department of Education Office of Postsecondary Education [-@EADA2018]. We analyzed the salaries of n_1 head coaches from NCAA Division I public schools, and n_2 head coaches from NCAA Division I private schools.

The data from public schools is/is not approximately normal with a sample mean of XXX and sample standard deviation of XXX. The data from public schools is/is not approximately normal with a sample mean of XXX and sample standard deviation of XXX. 

DESCRIBE MORE DATA HERE

<!-- %TODO Describe the data (1st paragraph) as well as analysis results (order of results should correspond to hypotheses listed)-->

## Describing and Visualizing Head Coach data
```{r hist of head coach, message=FALSE}
hd.coach.salary_data <- ncaa.div.i_data %>%
  select(sector_name, HDCOACH_SALARY_MEN)
hd.coach.salary_data %>%
  ggplot(aes(x = HDCOACH_SALARY_MEN, fill = sector_name)) +
  geom_histogram(position = "dodge") +
  labs(title="Histogram of NCAA Div. I Men's Head Coach Salary", x = "Salary (USD)",
       y = "Count", fill = "School Type:") +
  theme(
    legend.position="bottom",
    panel.background = element_rect(fill = 'white', color = 'grey'),
    panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                color = "grey"), 
    panel.grid.minor = element_line(size = 0.1, linetype = 'solid',
                                color = "grey")
  )
```

## Describing and Visualizing Participation of Men and Head Coach Salaries
```{r scatter plot of head coach salaries}
men.participation.data <- ncaa.div.i_data %>%
  select(IL_PARTIC_MEN, HDCOACH_SALARY_MEN)
men.participation.data %>% ggplot(aes(x=IL_PARTIC_MEN, y=HDCOACH_SALARY_MEN)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(title="Scatterplot of Men's Head Coach Salary by Men Particiation Counts", x = "Count of Men Participating",
       y = "Head Coach's Salary (USD)") +
  theme(
    panel.background = element_rect(fill = 'white', color = 'grey'),
    panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                color = "grey"), 
    panel.grid.minor = element_line(size = 0.1, linetype = 'solid',
                                color = "grey")
  )
```

## Describing and Visualizing Participation of Men and Head Coach Salaries
```{r scatter plot of head coach salaries}
men.participation.data <- ncaa.div.i_data %>%
  select(IL_PARTIC_MEN, HDCOACH_SALARY_MEN)
men.participation.data %>% ggplot(aes(x=IL_PARTIC_MEN, y=HDCOACH_SALARY_MEN)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(title="Scatterplot of Men's Head Coach Salary by Men Particiation Counts", x = "Count of Men Participating",
       y = "Head Coach's Salary (USD)") +
  theme(
    panel.background = element_rect(fill = 'white', color = 'grey'),
    panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                color = "grey"), 
    panel.grid.minor = element_line(size = 0.1, linetype = 'solid',
                                color = "grey")
  )
```

## Results of Head Coach data

## Results of Particiption of Men and Head Coach Salaries

# Conclusion
<!-- %TODO Re-state the results in terms of the larger picture; state limitations and opportunities for future research (20 points) -->

## Session Info
```{r Session Info}
sessionInfo()
```

# References
