---
title: "Stats Modeling Project"
author: "Group 5: Xin Jin, Reid Ginoza, Heidi Lovejoy"
date: "11/18/2019"
# header-includes:
#    - \usepackage{amsmath, amssymb}
   # - \usepackage{natbib}
   # - \usepackage{hyperref}
output: pdf_document
bibliography: bibliography.bib
nocite: | 
  @rlang
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pgirmess)
library(pander)
library(tidyverse)
library(knitr)
# library(kableExtra) # for tables
library(magrittr)  # provides more pipe oeprators
library(haven)
options(knitr.table.format = "latex")
```

```{r support functions}
almost_sas <- function(aov.results){
  par(mfrow=c(2,2))
  plot(aov.results, which=1)
  plot(aov.results, which=2)
  aov_residuals <- residuals(aov.results)
  plot(density(aov_residuals))
  hist(aov_residuals)
}
```

```{r}
p.value.string = function(p.value){
  p.value <- round(p.value, digits=4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}
```

```{r import data}
institution_data <- as_tibble(read_spss('InstLevel.sav')) %>%
  select(  # TODO will change later when we decide what variables we want
  -unitid, -addr1_txt, -addr2_txt, -city_txt, -zip_text, -sector_cd,
  -ClassificationCode, -ClassificationOther
  ) %>%
  filter(grepl('4-year', sector_name))  %>% # only 4-year Schools
  filter(is.na(IL_PARTIC_COED_MEN) | IL_PARTIC_COED_MEN == 0 ) %>% # only schools with no male participants in coed
  filter(is.na(IL_PARTIC_COED_WOMEN) | IL_PARTIC_COED_WOMEN == 0 ) %>% # only schools with no female participants in coed
  select(-contains("COED")) %>% # ignore variables with the word "coed"
  mutate(Grand.Total.Profits = GRND_TOTAL_REVENUE - GRND_TOTAL_EXPENSE) %>%
  filter(!is.na(HDCOACH_SALARY_MEN)) %>% # filter out NA
  filter(!is.na(HDCOACH_SALARY_WOMEN))  # filter out NA
ncaa.div.i_data <- institution_data %>%
  filter(grepl('NCAA Division I-', classification_name)) # only NCAA division 1
```

```{r Notes to self, include=FALSE}
##### All 4-year Institutions
# institution_data <- read_spss('InstLevel.sav') %>%
#   select(  # TODO will change later when we decide what variables we want
#   -unitid, -addr1_txt, -addr2_txt, -city_txt, -zip_text, -sector_cd,
#   -ClassificationCode, -ClassificationOther
#   ) %>%
#   filter(grepl('4-year', sector_name))  %>% # only 4-year Schools
#   filter(is.na(IL_PARTIC_COED_MEN) | IL_PARTIC_COED_MEN == 0 ) %>% # only schools with no male participants in coed
#   filter(is.na(IL_PARTIC_COED_WOMEN) | IL_PARTIC_COED_WOMEN == 0 ) %>% # only schools with no female participants in coed
#   select(-contains("COED")) # ignore variables with the word "coed"
##### To view the school data
# school_data <- read_spss('Schools.sav') %>% select(
#   -unitid, -addr1_txt, -addr2_txt, -zip_text, -ClassificationCode,-ClassificationOther,
#   -sector_cd
#   )
#####
# To view variables included as a list (although you could probably just view the data):
# institution_columns <- institution_data %>% colnames()
#####
# To view number of NAs in a column:
# count.of.na <- institution_data %>% summarise_all(~sum(is.na(.)))
#####
# To view unique entries of a column
# divisions <- institution_data$classification_name %>% unique()
```

# Introduction
<!-- %TODO Briefly give meaning to your project – what is the background and why is this project important?  (20 points) -->

This project examines potential(?) salary differences between head coaches from public and private NCAA Division I institutions. We used the data from the Equity in Athletics Survey, Year 2017-2018, from the U.S. Department of Education Office of Postsecondary Education [-@EADA2018].

# Hypotheses

## Divisions and Profits

## Head Coaches of Men's Teams and School Type
<!-- %TODO Clearly state the hypotheses being tested. (20 points) -->
<!-- Note: we need the two spaces at the end of the first line of the hypothesis test-->
\(H_0\):  
\(H_1\):


```{r}
wilcoxon.attempt <- institution_data %>%
  select(classification_name, HDCOACH_SALARY_MEN, HDCOACH_SALARY_WOMEN) %>%
  gather("Gender", "Salary", -classification_name)
wilcoxon.attempt$Gender %<>% recode(HDCOACH_SALARY_MEN="Men", HDCOACH_SALARY_WOMEN="Women")

wilcoxon.attempt 
```


```{r damage wilcoxon test}
wilcoxon.attempt_results <- wilcox.test(Salary ~ Gender, alternative = 'two.sided', data = wilcoxon.attempt, paired = FALSE)

wilcoxon.attempt_results


gender_wilcoxon_p.value <- wilcoxon.attempt_results$p.value %>% p.value.string
```

**Hypotheses**  
\(H_0\):  \(M_{male} = M_{female}\)  
\(H_1\): \(M_{male} \ne M_{female}\)

**Test Statistics**  
$W=`r wilcoxon.attempt_results$statistic`$

**$p$-value**  
$`r gender_wilcoxon_p.value`$

**Rejection Region**  
Reject $H_0$ if $p<\alpha$, where $\alpha = 0.05$.

**Conclusion**<br>


