---
title: "Old Notes"
author: "Group 5: Xin Jin, Reid Ginoza, Heidi Lovejoy"
date: "11/18/2019"
# header-includes:
#    - \usepackage{amsmath, amssymb}
   # - \usepackage{natbib}
   # - \usepackage{hyperref}
output: pdf_document
bibliography: bibliography.bib
nocite: | 
  @rlang
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pgirmess)
library(pander)
library(tidyverse)
library(knitr)
library(kableExtra) # for tables
library(magrittr)  # provides more pipe oeprators
library(haven)
options(knitr.table.format = "latex")

almost_sas <- function(aov.results){
  par(mfrow=c(2,2))
  plot(aov.results, which=1)
  plot(aov.results, which=2)
  aov_residuals <- residuals(aov.results)
  plot(density(aov_residuals))
  hist(aov_residuals)
}
```

```{r import data}
institution_data <- read_spss('InstLevel.sav') %>%
  select(institution_name, state_cd, ClassificationCode, classification_name, sector_cd, sector_name, HDCOACH_SALARY_MEN, HDCOACH_SALARY_WOMEN, IL_PARTIC_MEN, IL_PARTIC_WOMEN, IL_PARTIC_COED_MEN, IL_PARTIC_COED_WOMEN) %>%
  filter(grepl('4-year', sector_name))  %>% # only 4-year Schools
  filter(is.na(IL_PARTIC_COED_MEN) | IL_PARTIC_COED_MEN == 0 ) %>% # only schools with no male participants in coed
  filter(is.na(IL_PARTIC_COED_WOMEN) | IL_PARTIC_COED_WOMEN == 0 ) %>% # only schools with no female participants in coed
  select(-contains("COED")) %>% # ignore variables with the word "coed"
  filter(!is.na(HDCOACH_SALARY_MEN)) %>% # filter out NA
  filter(!is.na(HDCOACH_SALARY_WOMEN))  # filter out NA

ncaa.div.i_data <- institution_data %>%
  filter(grepl('NCAA Division I-', classification_name)) # only NCAA division 1

men_data <- institution_data %>%
  select(-contains("WOMEN")) %>% 
  rename(HD_SALARY = HDCOACH_SALARY_MEN, IL_PARTIC = IL_PARTIC_MEN)

women_data <- institution_data %>%
  select(-contains("_MEN")) %>%
  rename(HD_SALARY = HDCOACH_SALARY_WOMEN, IL_PARTIC = IL_PARTIC_WOMEN)

gender.separated_data <- bind_rows(
  men_data %>% add_column(Gender = "Men"),
  women_data %>% add_column(Gender = "Women")
)
```

# Introduction
<!-- %TODO Briefly give meaning to your project – what is the background and why is this project important?  (20 points) -->
This project examined predictors of head coach salaries of sports teams at four-year colleges and universities in the United States. We used the data from the Equity in Athletics Survey, Year 2017-2018, from the U.S. Department of Education Office of Postsecondary Education [-@EADA2018]. Our analysis focused on the differences of head coach salaries between male and female teams, of classification, and of the number of student participants at an institution level.

The full data set listed completed surveys for 2079 institutions. The 578 two-year colleges were excluded, as well as the 120 four-year schools with coed teams and 18 other institutions with missing values for head coach salaries. Our working data set had 1363 four-year institutions without coed sports teams.

# Hypotheses

## Divisions and Profits


## Head Coaches of Men's Teams and School Type
<!-- %TODO Clearly state the hypotheses being tested. (20 points) -->
<!-- Note: we need the two spaces at the end of the first line of the hypothesis test-->
\(H_0\):  
\(H_1\):

## Head Coaches of Men's Teams and Participation
<!-- %TODO Clearly state the hypotheses being tested. (20 points) -->
\(H_0\):  
\(H_1\):

## Difference in Classifications
There are 18 different classifications, including NCAA Division I-FCS, NCAA II without Football, and so forth. This question looks at the means of the profits for each classification.
\(H_0\): \(\mu_1 = \mu_2 = \dots = \mu_{18}\)  
\(H_1\): At least one is different

## Difference in Classfications (Nonparametric)

# Methods
<!-- %TODO State and explain the methodology used to test the hypotheses.  (20 points) -->
Data was analyzed using R version 3.6.1 (2019-07-05) [@rlang]. Data is described using counts (percentage) for categorical variables and mean (standard deviation) for continuous variables.

# Results
<!-- %TODO Describe the data (1st paragraph) -->

## Results of Difference in Classification
<!-- ```{r ANOVA for sectors} -->
<!-- classification.results <- aov(Grand.Total.Profits ~ classification_name, data = institution_data) -->
<!-- almost_sas(classification.results) -->
<!-- ``` -->

<!-- ```{r Classification Kruskal-Wallis} -->
<!-- classification_kruskal.results <- kruskal.test(Grand.Total.Profits ~ classification_name, data = institution_data) -->
<!-- classification_post.hoc <- kruskalmc(Grand.Total.Profits ~ classification_name, data = institution_data) -->

<!-- # pander(classification_post.hoc[[3]], style='rmarkdown')  # TODO: shorten table! -->
<!-- ``` -->

## Describing and Visualizing Head Coach data
```{r hist of head coach, message=FALSE}
hd.coach.salary_data <- ncaa.div.i_data %>%
  select(sector_name, HDCOACH_SALARY_MEN)

hd.coach.salary_data %>%
  ggplot(aes(x = HDCOACH_SALARY_MEN, fill = sector_name)) +
  geom_histogram(position = "dodge") +
  labs(title="Histogram of NCAA Div. I Men's Head Coach Salary", x = "Salary (USD)",
       y = "Count", fill = "School Type:") +
  theme(
    legend.position="bottom",
    panel.background = element_rect(fill = 'white', color = 'grey'),
    panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                color = "grey"), 
    panel.grid.minor = element_line(size = 0.1, linetype = 'solid',
                                color = "grey")
  )
```

```{r qq plot of Private Schools head coach}
hd.coach.salary_data %>%
  filter(grepl('Private', sector_name)) %>% 
  ggplot(aes(sample = HDCOACH_SALARY_MEN)) + 
  stat_qq() + stat_qq_line() + theme_bw()
```
```{r qq plot Public Schools head cach}
hd.coach.salary_data %>%
  filter(grepl('Public', sector_name)) %>% 
  ggplot(aes(sample = HDCOACH_SALARY_MEN)) + 
  stat_qq() + stat_qq_line() + theme_bw()
```

## Describing and Visualizing Participation of Men and Head Coach Salaries
```{r scatter plot of head coach salaries}
men.participation.data <- ncaa.div.i_data %>%
  select(IL_PARTIC_MEN, HDCOACH_SALARY_MEN)

men.participation.data %>% ggplot(aes(x=IL_PARTIC_MEN, y=HDCOACH_SALARY_MEN)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(title="Scatterplot of Men's Head Coach Salary by Men Particiation Counts", x = "Count of Men Participating",
       y = "Head Coach's Salary (USD)") +
  theme(
    panel.background = element_rect(fill = 'white', color = 'grey'),
    panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                color = "grey"), 
    panel.grid.minor = element_line(size = 0.1, linetype = 'solid',
                                color = "grey")
  )
```

<!-- %TODO And analysis results (order of results should correspond to hypotheses listed).  (20 points) -->
## Results of Head Coach data

## Results of Particiption of Men and Head Coach Salaries

# Conclusion
<!-- %TODO Restate the results in terms of the larger picture; state limitations and opportunities for future research (20 points) -->

Our data set contained only the averages of head coach salaries of a given institution, so we were unable to capture any variance within the institution.

## Session Info
```{r Session Info}
sessionInfo()
```

# References