---
title: "Chapter 7 Homework"
author: "Reid Ginoza"
date: "10/10/2019"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(EnvStats)
library(stats)
library(car)
library(knitr)
library(kableExtra)
```

```{r Support Functions}
### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits = 4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}
```

# Chapter 7 Homework

## 2. Data from Problem 7.20
```{r Import Tire Data}
tire_data <- as_tibble(read_csv("ex7-20.csv", col_names = TRUE, col_types = 'dd')) %>% gather('Brand', 'Distance')
```

### (a)  Find the mean and standard deviation of the data, split by brand.
```{r Tire Mean and SD}
tire_summary <- tire_data %>% group_by(Brand) %>% summarise(Mean = mean(Distance), SD = round(sd(Distance), digits = 3))
tire_summary %>% select(Brand, Mean, SD) %>% kable(align ='lrr',format.args = list(trim = TRUE)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                font_size = 14,
                position = "left")
```

The mean (standard deviation) of the distance of Brand I tires is `r tire_summary %>% filter(Brand == 'Brand I') %>% select(Mean)` (`r tire_summary %>% filter(Brand == 'Brand I') %>% select(SD)`) thousands of miles.<br>
The mean (standard deviation) of the distance of Brand II tires is `r tire_summary %>% filter(Brand == 'Brand II') %>% select(Mean)` (`r tire_summary %>% filter(Brand == 'Brand II') %>% select(SD)`) thousands of miles.

### (b)  Formally  test  to  determine  if  there  is  a  difference  between  the  standard  deviations  of  the  two brands at the α=0.01 level.

```{r Tires Hypothesis Test}
tires_h.test <- var.test((tire_data %>% filter(Brand == 'Brand I'))$Distance, (tire_data %>% filter(Brand == 'Brand II'))$Distance, alternative = "two.sided", conf.level = 0.99, ratio = 1)

tires_p.value <- tires_h.test$p.value %>% p.value.string
```

**Hypotheses**<br>
$H_0: \sigma_I^2 = \sigma_{II}^2$<br>
$H_1: \sigma_I^2 \ne \sigma_{II}^2$

**Test Statistic**<br>
$F_0=`r tires_h.test$statistic`$

**$p$-value**<br>
$`r tires_p.value`$

**Rejection Region**<br>
Reject $H_0$ if $p<\alpha$, where $\alpha = 0.01$.

**Conclusion**<br>
Reject $H_0$. There is sufficient evidence to suggest that the standard deviation of the Brand II tires is different from the standard deviation of Brand I tires.

### (c)  Construct the 99% confidence interval for $\frac{\sigma_I}{\sigma_{II}}$.
```{r Tires Confidence Interval}
tires_sd_conf.int <- round(sqrt(tires_h.test$conf.int), digits = 2)
```

The 99% confidence interval for the ratio $\frac{\sigma_I}{\sigma_{II}}$ is (`r tires_sd_conf.int`).

### (d)  If  we  were  to  go  on  and  test  the  means  of  the  two  brands,  would  we  use  the  pooled $t$-test  or Satterthwaite’s?  Why?
We would use Satterthwaite's approximation for the $t$-test, which allows for populations that do not have the same variance, and we just arrived at the conclusion that the two variances (or standard devations) are not equal.

## Session Info
```{r}
sessionInfo()
```