---
title: "Chapter 7 Homework"
author: "Reid Ginoza"
date: "10/10/2019"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(EnvStats)
library(stats)
library(car)
library(knitr)
library(kableExtra)
```

```{r Support Functions}
### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits = 4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}
```

# Chapter 7 Homework

## 1. Data from Problem 7.5
Amount of juice in 24 containers.

* $n=24$
* $\mu = 64.3$
* $\sigma_0 = 0.15$

```{r Import Juice Data}
juice_data <- as_tibble(read_csv("ex7-5.csv", col_names = TRUE, col_types = 'd'))
```

### (a) Find the mean and standard deviation of the data
```{r Juice Mean and SD}
mean_juice <- round(mean(juice_data$ounces), digits = 2)
sd_juice <- round(sd(juice_data$ounces), digits = 2)
```

The mean (standard deviation) of the amount of juice in a container is `r mean_juice` (`r sd_juice`) ounces.

### (b)  Formally test to determine if the process standard deviation is greater than 0.15 ounces at the α=0.10 level.

```{r Juice Hypothesis Test}
juice_h.test <- varTest(juice_data$ounces, alternative = "greater", conf.level = 0.90, sigma.squared = 0.15^2)
juice_p.value <- juice_h.test$p.value %>% p.value.string
```

**Hypotheses**<br>
$H_0: \sigma \le 0.15$<br>
$H_1: \sigma > 0.15$

**Test Statistic**<br>
$\chi^2_0 = `r round(juice_h.test$statistic, digits = 2)`$

**$p$-value**<br>
$`r juice_p.value`$

**Rejection Region**<br>
Reject $H_0$ if $p<\alpha$, where $\alpha = 0.10$.

**Conclusion**<br>
Reject $H_0$. There is sufficent evidence to suggest that there are too many underfilled and overfilled containers orange juice containers.

### (c)  Construct the 90% confidence interval for σ.
```{r Juice Confidence Interval}
juice_h.test_two.sided <- varTest(juice_data$ounces, alternative = "two.sided", conf.level = 0.90, sigma.squared = 0.15^2)
juice_ci_sd <- round(sqrt(juice_h.test_two.sided$conf.int), digits = 2)
```

The confidence interval for the orange juice process standard deviation $\sigma$ is (`r juice_ci_sd`).

## 2. Data from Problem 7.20
```{r Import Tire Data}
tire_data <- as_tibble(read_csv("ex7-20.csv", col_names = TRUE, col_types = 'dd')) %>% gather('Brand', 'Distance')
```

### (a)  Find the mean and standard deviation of the data, split by brand.
```{r Tire Mean and SD}
tire_summary <- tire_data %>% group_by(Brand) %>% summarise(Mean = mean(Distance), SD = round(sd(Distance), digits = 3))
tire_summary %>% select(Brand, Mean, SD) %>% kable(align ='lrr',format.args = list(trim = TRUE)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                font_size = 14,
                position = "left")
```

The mean (standard deviation) of the distance of Brand I tires is `r tire_summary %>% filter(Brand == 'Brand I') %>% select(Mean)` (`r tire_summary %>% filter(Brand == 'Brand I') %>% select(SD)`) thousands of miles.<br>
The mean (standard deviation) of the distance of Brand II tires is `r tire_summary %>% filter(Brand == 'Brand II') %>% select(Mean)` (`r tire_summary %>% filter(Brand == 'Brand II') %>% select(SD)`) thousands of miles.

### (b)  Formally  test  to  determine  if  there  is  a  difference  between  the  standard  deviations  of  the  two brands at the α=0.01 level.

```{r Tires Hypothesis Test}
tires_h.test <- var.test((tire_data %>% filter(Brand == 'Brand I'))$Distance, (tire_data %>% filter(Brand == 'Brand II'))$Distance, alternative = "two.sided", conf.level = 0.99, ratio = 1)

tires_p.value <- tires_h.test$p.value %>% p.value.string
```

**Hypotheses**<br>
$H_0: \sigma_I = \sigma_{II}$<br>
$H_1: \sigma_I \ne \sigma_{II}$

**Test Statistic**<br>
$F_0=`r tires_h.test$statistic`$

**$p$-value**<br>
$`r tires_p.value`$

**Rejection Region**<br>
Reject $H_0$ if $p<\alpha$, where $\alpha = 0.01$.

**Conclusion**<br>
Reject $H_0$. There is sufficient evidence to suggest that the standard deviation of the Brand II tires is different from the standard deviation of Brand I tires.

### (c)  Construct the 99% confidence interval for $\frac{\sigma_I}{\sigma_{II}}$.
```{r Tires Confidence Interval}
tires_sd_conf.int <- round(sqrt(tires_h.test$conf.int), digits = 2)
```

The 99% confidence interval for the ratio $\frac{\sigma_I}{\sigma_{II}}$ is (`r tires_sd_conf.int`).

### (d)  If  we  were  to  go  on  and  test  the  means  of  the  two  brands,  would  we  use  the  pooled $t$-test  or Satterthwaite’s?  Why?
We would use Satterthwaite's approximation for the $t$-test, which allows for populations that do not have the same variance/standard deviation, and we just arrived at the conclusion that the two population standard deviations are not equal.

## 3. Data from Problem 7.29
```{r Import Oven Data}
oven_data <- as_tibble(read_csv("ex7-29.csv", col_names = TRUE, col_types = 'ddd')) %>% gather('Oven', 'Temperature')
```

### (a)  Find the mean and standard deviation, split by oven.
```{r Oven Mean and SD}
oven_summary <- oven_data %>% group_by(Oven) %>% summarize(Mean = round(mean(Temperature), digits = 2), SD = round(sd(Temperature), digits = 2))
oven_summary %>% select(Oven, Mean, SD) %>% kable(align ='lrr',format.args = list(trim = TRUE)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                font_size = 14,
                position = "left")
```

The mean (standard deviation) temperature of Oven I is  `r oven_summary %>% filter(Oven == 'Oven1') %>% select(Mean)` (`r oven_summary %>% filter(Oven == 'Oven1') %>% select(SD)`).<br>
The mean (standard deviation) temperature of Oven II is  `r oven_summary %>% filter(Oven == 'Oven2') %>% select(Mean)` (`r oven_summary %>% filter(Oven == 'Oven2') %>% select(SD)`).<br>
The mean (standard deviation) temperature of Oven III is  `r oven_summary %>% filter(Oven == 'Oven3') %>% select(Mean)` (`r oven_summary %>% filter(Oven == 'Oven3') %>% select(SD)`).

### (b)  Formally test to determine if there is a difference between the standard deviations at the α=0.05 level.
```{r Over Hypothesis Test}
oven_h.test <- leveneTest(Temperature ~ Oven, data = oven_data, center = median)
oven_p.value <- p.value.string(oven_h.test$"Pr(>F)"[1])
```

**Hypotheses**<br>
$H_0: \sigma_I = \sigma_{II} = \sigma_{III}$<br>
$H_1:$ At least one is different

**Test Statistic**<br>
$F_0 = `r oven_h.test$F[1]`$

**$p$-value**<br>
$`r oven_p.value`$

**Rejection Region**<br>
Reject $H_0$ if $p<\alpha$, where $\alpha=0.05$.

**Conclusion**<br>
Reject $H_0$. There is sufficient evidence to suggest that at least one of the standard deviations is different from the others.

### (c)  Find the 95% CI for the σ of each oven.
```{r Oven Individual CIs}
oven1_ci <- round(sqrt(varTest((oven_data %>% filter(Oven == 'Oven1'))$Temperature, alternative = "two.sided", conf.level = 0.95)$conf.int), digits = 2)
oven2_ci <- round(sqrt(varTest((oven_data %>% filter(Oven == 'Oven2'))$Temperature, alternative = "two.sided", conf.level = 0.95)$conf.int), digits = 2)
oven3_ci <- round(sqrt(varTest((oven_data %>% filter(Oven == 'Oven3'))$Temperature, alternative = "two.sided", conf.level = 0.95)$conf.int), digits = 2)
```

| Oven      | 95% Confidence Interval for $\sigma$|
| ----------- | ----------- |
| Oven 1  | $(`r oven1_ci`)$ |
| Oven 2  | $(`r oven2_ci`)$ |
| Oven 3  | $(`r oven3_ci`)$ |


### (d)  Do the CIs in part (c) match your hypothesis test results in part (a)?  Why or why not?
Yes, my confidence intervals agree with my results. Oven 3's interval does not overlap with either of the other two, so it makes sense the 

### (e)  Do you see any potential issues with making the comparison requested in part (d)?
Not for this problem. Even before the 95% confidence intervals for the standard deviations were calculated, the sample standard deviation for Oven 3 was much higher than that of Oven 1 and Oven 2. Perhaps if the Oven 3 confidence interval overlapped with that of Oven 2, there would be a concern.

## Session Info
```{r}
sessionInfo()
```