---
title: "Chapter 7 Homework"
author: "Reid Ginoza"
date: "10/10/2019"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(EnvStats)
library(stats)
library(car)
library(knitr)
library(kableExtra)
```

```{r Support Functions}
### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits = 4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}
```

# Chapter 7 Homework
## 1. Data from Problem 7.5
Amount of juice in 24 containers.

* $n=24$
* $\mu = 64.3$
* $\sigma_0 = 0.15$

```{r Import Juice Data}
juice_data <- as_tibble(read_csv("ex7-5.csv", col_names = TRUE, col_types = 'd'))
```

### (a) Find the mean and standard deviation of the data
```{r Juice Mean and SD}
mean_juice <- round(mean(juice_data$ounces), digits = 2)
sd_juice <- round(sd(juice_data$ounces), digits = 2)
```

The mean (standard deviation) of amount of juice is `r mean_juice` (`r sd_juice`) ounces.

### (b)  Formally test to determine if the process standard deviation is greater than 0.15 ounces at the α=0.10 level.

```{r Juice Hypothesis Test}
juice_h.test <- varTest(juice_data$ounces, alternative = "greater", conf.level = 0.90, sigma.squared = 0.15^2)
juice_p.value <- juice_h.test$p.value %>% p.value.string
```

**Hypotheses**<br>
$H_0: \sigma \le 0.15$<br>
$H_1: \sigma > 0.15$

**Test Statistic**<br>
$\chi^2_0 = `r round(juice_h.test$statistic, digits = 2)`$

**$p$-value**<br>
$`r juice_p.value`$

**Rejection Region**<br>
Reject $H_0$ if $p<\alpha$, where $\alpha = 0.10$.

**Conclusion**<br>
Reject $H_0$. There is sufficent evidence to suggest that there are have too many underfilled and overfilled containers orange juice containers.

### (c)  Construct the 90% confidence interval for σ.
```{r Juice Confidence Interval}
juice_h.test_two.sided <- varTest(juice_data$ounces, alternative = "two.sided", conf.level = 0.90, sigma.squared = 0.15^2)
juice_ci_sd <- round(sqrt(juice_h.test_two.sided$conf.int), digits = 2)
```

The confidence interval for the orange juice process standard deviation $\sigma$ is (`r juice_ci_sd`).

## Session Info
```{r}
sessionInfo()
```