---
title: "Chapter 7 Homework"
author: "Reid Ginoza"
date: "10/10/2019"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(EnvStats)
library(stats)
library(car)
library(knitr)
library(kableExtra)
```

```{r Support Functions}
### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits = 4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}
```

# Chapter 7 Homework

## 3. Data from Problem 7.29
```{r Import Oven Data}
oven_data <- as_tibble(read_csv("ex7-29.csv", col_names = TRUE, col_types = 'ddd')) %>% gather('Oven', 'Temperature')
```

### (a)  Find the mean and standard deviation, split by oven.
```{r Oven Mean and SD}
oven_summary <- oven_data %>% group_by(Oven) %>% summarize(Mean = round(mean(Temperature), digits = 2), SD = round(sd(Temperature), digits = 2))
oven_summary %>% select(Oven, Mean, SD) %>% kable(align ='lrr',format.args = list(trim = TRUE)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                font_size = 14,
                position = "left")
```

The mean (standard deviation) temperature of Oven I is  `r oven_summary %>% filter(Oven == 'Oven1') %>% select(Mean)` (`r oven_summary %>% filter(Oven == 'Oven1') %>% select(SD)`).<br>
The mean (standard deviation) temperature of Oven II is  `r oven_summary %>% filter(Oven == 'Oven2') %>% select(Mean)` (`r oven_summary %>% filter(Oven == 'Oven2') %>% select(SD)`).<br>
The mean (standard deviation) temperature of Oven III is  `r oven_summary %>% filter(Oven == 'Oven3') %>% select(Mean)` (`r oven_summary %>% filter(Oven == 'Oven3') %>% select(SD)`).

### (b)  Formally test to determine if there is a difference between the standard deviations at the α=0.05 level.
```{r Over Hypothesis Test}
oven_h.test <- leveneTest(Temperature ~ Oven, data = oven_data, center = median)
oven_p.value <- p.value.string(oven_h.test$"Pr(>F)"[1])
```

**Hypotheses**<br>
$H_0: \sigma_I = \sigma_{II} = \sigma_{III}$<br>
$H_1:$ At least one is different

**Test Statistic**<br>
$F_0 = `r oven_h.test$F[1]`$

**$p$-value**<br>
$`r oven_p.value`$

**Rejection Region**<br>
Reject $H_0$ if $p<\alpha$, where $\alpha=0.05$.

**Conclusion**<br>
Reject $H_0$. There is sufficient evidence to suggest that at least one of the standard deviations is different from the others.

### (c)  Find the 95% CI for the σ of each oven.
```{r Oven Individual CIs}
oven1_ci <- round(sqrt(varTest((oven_data %>% filter(Oven == 'Oven1'))$Temperature, alternative = "two.sided", conf.level = 0.95)$conf.int), digits = 2)
oven2_ci <- round(sqrt(varTest((oven_data %>% filter(Oven == 'Oven2'))$Temperature, alternative = "two.sided", conf.level = 0.95)$conf.int), digits = 2)
oven3_ci <- round(sqrt(varTest((oven_data %>% filter(Oven == 'Oven3'))$Temperature, alternative = "two.sided", conf.level = 0.95)$conf.int), digits = 2)
```

| Oven      | 95% Confidence Interval for $\sigma$|
| ----------- | ----------- |
| Oven 1  | $(`r oven1_ci`)$ |
| Oven 2  | $(`r oven2_ci`)$ |
| Oven 3  | $(`r oven3_ci`)$ |


### (d)  Do the CIs in part (c) match your hypothesis test results in part (a)?  Why or why not?
Yes, my confidence intervals agree with my results. Oven 3's interval does not overlap with either of the other two, so it makes sense the 

### (e)  Do you see any potential issues with making the comparison requested in part (d)?
Not for this problem. Even before the 95% confidence intervals for the standard deviations were calculated, the sample standard deviation for Oven 3 was much higher than that of Oven 1 and Oven 2. Perhaps if the Oven 3 confidence interval overlapped with that of Oven 2, there would be a concern.

## Session Info
```{r}
sessionInfo()
```